A search for intermediate mass black holes mergers in the \\second LIGO--Virgo observing run with the Bayes Coherence Ratio

### 2021-06-07 main.tex
incorporating erics suggestions

\title{A Bayesian ranking-statistic for binary black holes mergers in the second LIGO observing run}

@@

\begin{abstract}
The detection of an intermediate-mass black hole population ($10^2-10^6$~\msun) will provide probes to their formation environments (e.g., disks of active galactic nuclei, globular clusters) and illuminate a potential pathway to produce supermassive black holes. Numerous methods to search for $10^4-10^6\ \msun$ intermediate-mass black holes exist, while the primary method to find $10^2-10^3\ \msun$ is by identifying gravitational-waves emitted from the merger of stellar binaries. Ground-based gravitational-wave detectors are in principle sensitive to such mergers and have found one $142^{+28}_{-16}\ \msun$ intermediate-mass black hole. In reality, ground-based detector data contain numerous short-duration instrumental transients that mimic the gravitational-wave signals from mergers capable of producing intermediate-mass black holes, making intermediate-mass merger detection challenging. Here we demonstrate a ranking statistic utilizing Bayesian inference to detect binary black hole mergers with a total mass $>55$~\msun. We apply this technique to candidate events with total masses $>55$~\msun during LIGO's second observing run to search for previously unresolved gravitational-wave signals from this heavy category of binary black holes. Our analysis does not discover new intermediate-mass black holes. However, we find support for some stellar-mass binary black holes unreported in GWTC-1.
\end{abstract}

@@

%%%%%%---SECTIONS---%%%%%%%%%%%%
\section{Introduction}
Since the 1970s, there has been a steady accumulation of evidence for stellar mass ($M_\text{BH} < 10^{2} \ \msun$) and supermassive black holes ($M_\text{BH} > 10^{6} \ \msun$)~\cite{Webster:1972:Natur, Balick:1974:ApJ, Ghez:1998:ApJ, Genzel:2010:RvMP, Abbott:2019:PhRvX, EventHorizonTelescopeCollaboration:2019:ApJL, Abbott:2020:arXiv}.  However, there is a deficiency of observational evidence for black holes in the `intermediate-mass' range $10^{2} - 10^{6}\ \msun$. The discovery of an intermediate-mass black hole population will bridge this observational gap, probe intermediate-mass formation environments (e.g. accretion disks of active galactic nuclei~\cite{Tagawa:2021:ApJ, Li:2021:arXiv, Samsing:2020:arXiv, Tagawa:2020:ApJ, Ishibashi:2020:A&A, Grobner:2020:A&A, Yang:2019:PhRvL, McKernan:2019:ApJL, Yang:2019:ApJ, McKernan:2018:ApJ, Bellovary:2016:ApJL, McKernan:2014:MNRAS, McKernan:2012:MNRAS}, the centers of dense stellar clusters~\cite{Banerjee:2021:MNRASa, Zevin:2021:ApJ,Mapelli:2021:arXiv,Weatherford:2021:ApJL, Bouffanais:2021:arXiv, Ballone:2021:MNRAS, Kumamoto:2021:arXiv, Banerjee:2021:MNRASb, Martinez:2020:ApJ, Romero-Shaw:2020:ApJL, Anagnostou:2020:PASA}, Population-III stars~\cite{Toubiana:2021:PhRvL, Farrell:2021:MNRAS, Safarzadeh:2020:ApJL, Liu:2020:MNRAS, Inayoshi:2017:MNRAS}), and illuminate our understanding of supermassive black hole formation~\cite{Askar:2021:MNRAS, ArcaSedda:2019:arXiv, Amaro-Seoane:2007:CQGra, Gurkan:2006:ApJL}. A variety of techniques have been employed to search for $10^{4} - 10^{6}\ \msun$ intermediate mass candidates (reverberation mapping~\cite{Peterson:2014:SSRv}, direct kinematic measurements~\cite{Schodel:2002:Natur, Kiziltan:2017:Natur}, applying macroscopic galaxy to black hole mass scaling relations ($M_{BH}$-$\sigma$ and $M_{BH}$-L relations)~\cite{Graham:2013:ApJ, Wevers:2017:MNRAS}, studying  X-ray luminosity and spectra~\cite{Greene:2004:ApJ, Lin:2020:ApJL}, gravitational lensing of light curves~\cite{paynter_evidence_2021}, and others~\cite{Greene:2020:ARA&A, Koliopanos:2017:mbhe, Mezcua:2017:IJMPD})). However, these observational techniques are challenging to use for intermediate mass black holes due to their small sphere of influence compared to super-massive black holes~\cite{Mezcua:2017:IJMPD}. Additionally, some candidates discovered with these techniques can be attributed to sources without intermediate-mass black holes (e.g. clusters of stellar-mass black holes~\cite{Ridolfi:2016:MNRAS, Freire:2017:MNRAS}, anisotropic emission from neutron stars~\cite{Israel:2017:MNRAS, RodriguezCastillo:2020:ApJ}), while others have high uncertainties intrinsic to the observational techniques~\cite{Greene:2020:ARA&A}.

Compact binaries coalescences (CBCs) can provide unambiguous gravitational-wave signals for intermediate-mass candidates (e.g. the $142^{+28}_{-16}\ \msun$ remnant observed from the gravitational wave event GW190521~\cite{Abbott:2020:PhRvL}). As a binary's total mass $\text{M}_{T}$ is associated with its gravitational-wave merger frequency, $f\sim \text{M}_{T}^{-1}$,  ground based gravitational wave detectors ($f\sim 10^1 - 10^3\ \text{Hz}$) are sensitive to the last milliseconds of merging systems with $100 \msun < \text{M}_{T} < 400 \msun$~\cite{LIGOScientificCollaboration:2015:CQGra, Martynov:2016:PhRvD, Moore_2014}, while space-based detectors ($f \sim 10^{-2} - 10^1\ \text{Hz}$) can study the full signals of merging systems with $10^4 \msun < \text{M}_{T} < 10^7 \msun$~\cite{ Moore_2014, Lu:2019:PhRvD}. Because of the short-duration of intermediate-mass gravitational-wave signals in ground-based detector data streams, handling data quality is critical to their detection. The low-frequency ranges of ground-based observatories data-streams have numerous non-stationary terrestrial artifacts, called \textit{glitches}~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. Like signals from intermediate mass mergers, some glitches last for a fraction of a second, making them difficult to distinguish from the signals. These glitches that mimic astrophysical signals can severely decrease the confidence in detection of true gravitational-waves from black hole mergers with total masses $>55$~\msun ~\cite{pycbc_short_duration_transients}.

Although a significant fraction of the glitches can be discriminated by testing them for coherence amongst various ground based detectors and performing matched-filtering, these methods are insufficient to identify all glitches~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. One method to discriminate more glitches while searching for CBC gravitational-waves is to rely on a Bayesian odds~\cite{bci, kanner2016leveraging, BCR1, BCR2, bcr_gw151216, bayesian_odds}. Utilising Bayesian odds is the optimal method for discriminating glitches from signals~\cite{BCR2}. In this paper, we rank O2's candidate gravitational-wave signals from $55-500\ \msun$ total mass systems using a Bayesian method called the Bayesian Coherence Ratio \bcr~\cite{BCR1}. The \bcr used in this study is a bootstrap-Bayesian odds that estimates the true Bayesian odds. The \bcr is calculated using the probability of data under the hypothesis that the data contains coherent signals vs. incoherent glitches and empirically calculated prior-odds that describe how likely each hypothesis is. After computing the candidate \bcr values, we calculate the candidate's probability of astrophysical origins, \pastrobcr. Finally, if the candidate has a \pastro reported by the the LIGO-Virgo-KAGRA (LVK) collaboration in \GWTC~\cite{GWTC1}, the \pycbc-team~\cite{pycbc_code, pycbc_og0, pycbc_og1, pycbc_og2, pycbc_og3, pycbc_og4, pycbc_og5, pycbc_og6, pycbc_single_det, pycbc_ogc_2}, by the Institute of Advanced study's team (\IAS)~\cite{IAS0, IAS1, IAS2}, or by \citet{bayesian_odds}, we compare their \pastro with the \pastrobcr.  We highlight several innovations unique to this work. First, we use our Bayesian framework to analyze all candidate signals and background triggers (incoherent signal-free data) with total masses above $> 55 \msun$ in O2. Second, we empirically tune prior-odds for finding a signal or a glitch in O2 using \bcr distributions for background triggers and simulated signals. Finally,  we use the tuned-\bcr background distribution to compute how significant candidate signals are from the background.

We find that (a) high-mass events reported in the \GWTC, including GW170729 (the heaviest event in \GWTC) are very statistically significant $\pastrobcr>0.9$; (b) three out of the eight \IAS events and candidates have $\pastrobcr>0.5$, corroborating \IAS's detection claims for those events (GW170304, GW170727, GW170817A); and that (c) our ranking statistic does not identify any new intermediate-mass black holes, but does identify an unreported marginal stellar-mass binary black hole candidate, 170222 with $\pastrobcr\sim0.5$ %\avi{Need to doublecheck that this wasnt in the PyCBC sub-threshold paper}.

@@

The standard framework to identify CBC gravitational-wave signals in data is to quantify the significance of candidates with null-hypothesis significance testing~\cite{GWTC1, GWTC2}. In this framework, the candidates' ranking statistic is compared against a background distribution. The independent matched-filter searches, e.g. \pycbc~\cite{pycbc_og4}, \spiir~\cite{spiir} and \gstlal~\cite{sachdev2019gstlal}, and the coherent burst search \cwb~\cite{cwb} used by LVK to search for signals in gravitational-wave data all use ranking statistics in such a manner~\cite{GWTC1}. Both \pycbc and \gstlal's ranking statistic incorporate information of the relative likelihood that the data contains a coherent signal versus just noise. In contrast, \cwb's ranking statistic uses the information of coherent energy present in the network of detectors~\cite{GWTC1}.

Bayesian inference offers an alternative means to rank the significance of candidate events by computing the odds that the data contain a transient gravitational-wave signal vs. instrumental glitches or noise~\cite{BCR1}. This method relies on accurate models for the signal and glitch morphologies and populations~\cite{BCR1}. In principle, Bayesian odds is the optimal method for hypothesis testing~\cite{BCR2}. Much of its power comes from the Bayesian ``evidence'': a marginalized likelihood. The evidence is the correct likelihood of having obtained the data given a hypothesis. However, due to the high computational cost to calculate the evidence, it is not explicitly used in current matched filter searches. Here, we explore a middle-ground hybrid between frequentist matched filtered searches and Bayesian odds. We compute Bayesian evidences for foreground and background data under the assumption that they either contain coherent gravitational-wave signals, noise, or glitches. However, instead of computing true Bayesian odds, we use the evidences to form a bootstrapped distribution for the odds of foreground and background events to form a frequentist ranking statistic.

\subsection{The Bayesian Coherence Odds-Ratio}

@@

When $\mathcal{H}_S$ and $\mathcal{H}_I$ are precisely described and the prior-odds represent our true beliefs, the \bcrodds is a Bayesian odds-ratio. As an odds-ratio, the \bcrodds is the optimal discriminator between coherent signals and incoherent instrumental features. However, as the priors-odds are unknown, we \textit{tune} values for prior-odds, $\hat{\pi}^S$ and $\hat{\pi}^G$, to estimate \bcrodds with a ranking statistic, \bcr, given by
\begin{equation}
\label{eq:bcr}
\bcr = \frac{\hat{\pi}^S Z^S}{\prod\limits^D_{i=1} \ [\hat{\pi}^G Z^G_i + (1-\hat{\pi}^G)Z^N_i]}\ .
\end{equation}

In the limit where the estimated prior-odds equal the true prior-odds, $\bcr\to\bcrodds$. However, as we are uncertain what the true prior-odds are, it is invalid to use the \bcr as an odds-ratio to make an informed decision about whether a candidate is from an astrophysical or terrestrial source. Instead of interpreting the \bcr as a Bayesian odds-ratio, it can be used as a ranking statistic. Using the \bcr as a ranking statistic we can obtain a frequentist significance of a candidate \bcr-value measured against a background \bcr distribution.

@@

\subsection{Estimation of astrophysical signal probability}
Candidate \bcr-values are either statistically insignificant compared to the background \bcr distribution, implying the candidate is more probable to be an incoherent instrumental feature (the $\mathcal{H}_I$ null-hypothesis), or statistically significant relative to the background distribution, indicating the possible presence of an astrophysical signal (the $\mathcal{H}_S$ hypothesis). A false alarm probability with trial factors, \fap, for the candidate \bcr-value can quantify the significance. The \fap for a \bcr-value is the probability that a candidate originating from a non-astrophysical source can be falsely identified as a signal once in $N$ trials, and is given by 
\begin{equation}
    \fap = 1 - (1-f)^N \ ,
\end{equation}
where $f$ is the probability of observing a background $\bcr'$ greater than or equal to the candidate \bcr,
\begin{equation}
    f = \frac{\text{Count of } \bcr' \leq \bcr}{\text{Count of } \bcr'} \ .
\end{equation}

@@

\subsection{Data for Analysis}

To rank candidate event \bcr-values we need to compute a background \bcr distribution. We can obtain the background data needed from the LIGO Scientific collaboration's search pipelines that scan for gravitational-waves from compact binary mergers such as \gstlal~\cite{sachdev2019gstlal}, \mbta~\cite{MBTA}, \spiir~\cite{spiir} and \pycbc~\cite{pycbc_og4}. The output of \pycbc's search is a list of times and their corresponding \pycbc ranking statistic \pycbcstat values. Whenever a local maximum of $\pycbcstat > \snr_\text{T}$, where $\snr_\text{T}$ is some predetermined threshold value, the \pycbc search pipeline produces a single-detector \textit{trigger} associated with the detector and time $t_c$ where the apparent signal in the data has its merger~\cite{pycbc_og6}. \pycbc produces three categories of triggers: 
\begin{itemize}
    \item \textit{Candidate event trigger}: a trigger observed with coincident parameters amongst a network of detectors. 
    \item \textit{Simulated trigger}: a trigger detected from an artificially manufactured signal injected into detector data.  
    \item \textit{Background trigger}: a trigger obtained from incoherent signal-free background data (data manufactured by applying relative offsets, or time-slides, between the data of different detectors~\cite{pycbc_og6}). 
\end{itemize}

After obtaining these triggers from \pycbc, we can calculate the trigger's \bcr-values and use them to measure candidate triggers' statistical significance. The \bcr can be a powerful ranking statistic as it incorporates information of not only all possible binary black hole systems that might have merged to produce the trigger but also the various incoherent glitches that might cause a false-detection~\cite{BCR1}.

% \avi{Should I mention that we need \pycbc  as a pre-selection step to computing? }

@@

The most computationally intensive step during Bayesian inference is evaluating the likelihood $\mathcal{L}(d_i|\template)$. To accelerate our analysis, we use a likelihood that explicitly marginalizes over coalescence time, phase at coalescence, and luminosity distance (Eq.~80 from~\citet{intro_to_gw_bayes}).

@@

\subsection{Assigning \pastro to candidate events}
After calculating the \bcr for the entire set of background and simulated triggers, we calculate the background and simulated \bcr probability distributions for each 2-week time-frame of O2 data. These distributions are used to `tune' prior-odd $\hat{\pi}^S$ and $\hat{\pi}^G$ values as described in Appendix~\ref{apdx:tuning-prior-odds}.

@@

\input{Data/results_table}
We analyze the $60,996$ background, $5,146$ simulated, and $25$ candidate triggers reported by \pycbc's search on the data from LIGO's second observing run, restricting our analysis to the triggers that fall within our mass-space as described in Section~\ref{sec:method}.We also analyze events and candidate events reported by \GWTC and the \IAS group (note: some of these were identified as candidates by the \pycbc search). Table~\ref{tab:results} summaries the \pastrobcr, along with the \pastro of other pipelines for comparison. Although the various pipeline \pastro are not mathematically equivalent, by comparing pipeline \pastro values for a given candidate, we can compare how significant each pipeline deems various candidates. The $\hat{\pi}^S$ and $\hat{\pi}^G$ values utilized for each time-frame are reported in Appendix~\ref{apdx:alphabeta}.

@@

Until 2016, a majority of intermediate-mass black hole candidates were from electromagnetic observations. The dawn of gravitational wave astronomy has provided a new lens to identify and study these massive objects -- from the mergers of heavy black holes' gravitational wave emission. In the future, space-based gravitational wave detectors will uncover black holes in the upper end of the intermediate-mass range. In the present, ground-based gravitational-wave observatories can probe the lower-end of the intermediate-mass range. However, many short-duration terrestrial artifacts that mimic intermediate-mass gravitational-wave signals plague ground-based detectors, making intermediate-mass detection challenging.

@@

%%%%%%---ACKNOWLEDGMENTS---%%%%%%%%%%%%
\begin{acknowledgments}

The author gratefully thank the \pycbc team for providing the gravitational-wave foreground, background and simulated triggers from \pycbc's search of O2's data. We also warmly thank Ian Harry and Thomas Dent for answering questions about the \pycbc search's data products.

>>>>

\title{A search for intermediate mass black holes mergers in the \\second LIGO--Virgo observing run with the Bayes Coherence Ratio}

@@

\begin{abstract}
The detection of an intermediate-mass black hole population ($10^2-10^6$~\msun) will provide clues to their formation environments (e.g., disks of active galactic nuclei, globular clusters) and illuminate a potential pathway to produce supermassive black holes. Ground-based gravitational-wave detectors are in principle sensitive to such mergers and have been used to detect one $142^{+28}_{-16}\ \msun$ intermediate-mass black hole formation event. Ground-based detector data contain numerous short-duration noise transients that can mimic the gravitational-wave signals from merging intermediate-mass black holes, limiting the sensitivity of searches. Here we demonstrate a Bayesian-inspired ranking statistic to detect binary black hole mergers with a total mass $\gtrsim55$~\msun. We use this statistic to identify candidate events with total masses $>55$~\msun using data from LIGO's second observing run. Our analysis does not yield evidence for new intermediate-mass black holes. However, we find support for some stellar-mass binary black holes not reported in the first LIGO--Virgo gravitational-wave transient catalog, GWTC-1.
\end{abstract}

@@

%%%%%%---SECTIONS---%%%%%%%%%%%%
\section{Introduction}
Since the 1970s, there has been a steady accumulation of evidence for stellar mass ($M_\text{BH} < 10^{2} \ \msun$) and supermassive black holes ($M_\text{BH} > 10^{6} \ \msun$)~\cite{Webster:1972:Natur, Balick:1974:ApJ, Ghez:1998:ApJ, Genzel:2010:RvMP, Abbott:2019:PhRvX, EventHorizonTelescopeCollaboration:2019:ApJL, Abbott:2020:arXiv}.  However, there is a deficiency of observational evidence for black holes in the intermediate-mass range $10^{2} - 10^{6}\ \msun$. The discovery of an intermediate-mass black hole population will bridge this observational gap, probe intermediate-mass formation environments (e.g. accretion disks of active galactic nuclei~\cite{Tagawa:2021:ApJ, Li:2021:arXiv, Samsing:2020:arXiv, Tagawa:2020:ApJ, Ishibashi:2020:A&A, Grobner:2020:A&A, Yang:2019:PhRvL, McKernan:2019:ApJL, Yang:2019:ApJ, McKernan:2018:ApJ, Bellovary:2016:ApJL, McKernan:2014:MNRAS, McKernan:2012:MNRAS}, the centers of dense stellar clusters~\cite{Banerjee:2021:MNRASa, Zevin:2021:ApJ,Mapelli:2021:arXiv,Weatherford:2021:ApJL, Bouffanais:2021:arXiv, Ballone:2021:MNRAS, Kumamoto:2021:arXiv, Banerjee:2021:MNRASb, Martinez:2020:ApJ, Romero-Shaw:2020:ApJL, Anagnostou:2020:PASA}, Population-III stars~\cite{Toubiana:2021:PhRvL, Farrell:2021:MNRAS, Safarzadeh:2020:ApJL, Liu:2020:MNRAS, Inayoshi:2017:MNRAS}), and illuminate our understanding of supermassive black hole formation~\cite{Askar:2021:MNRAS, ArcaSedda:2019:arXiv, Amaro-Seoane:2007:CQGra, Gurkan:2006:ApJL}.

A variety of techniques have been employed to search for $10^{4} - 10^{6}\ \msun$ intermediate mass candidates including reverberation mapping~\cite{Peterson:2014:SSRv}, direct kinematic measurements~\cite{Schodel:2002:Natur, Kiziltan:2017:Natur}, applying macroscopic galaxy to black hole mass scaling relations ($M_{BH}$-$\sigma$ and $M_{BH}$-L relations)~\cite{Graham:2013:ApJ, Wevers:2017:MNRAS}, studying  X-ray luminosity and spectra~\cite{Greene:2004:ApJ, Lin:2020:ApJL}, gravitational lensing of gamma-ray burst light curves~\cite{paynter_evidence_2021}, and others~\cite{Greene:2020:ARA&A, Koliopanos:2017:mbhe, Mezcua:2017:IJMPD}). However, these observational techniques are challenging to use for intermediate mass black holes due to their small sphere of influence compared to super-massive black holes~\cite{Mezcua:2017:IJMPD}. Additionally, some IMBH candidates can be attributed to sources other than intermediate-mass black holes (e.g., clusters of stellar-mass black holes~\cite{Ridolfi:2016:MNRAS, Freire:2017:MNRAS}, anisotropic emission from neutron stars~\cite{Israel:2017:MNRAS, RodriguezCastillo:2020:ApJ}), while others have high uncertainties intrinsic to the observational techniques~\cite{Greene:2020:ARA&A}. 
\et{Consider replacing previous sentence that says none of the current IMBH candidates are widely accepted within the astronomical community as unambiguous.}

Compact binary coalescences (CBCs) can provide unambiguous gravitational-wave signals for intermediate-mass candidates (e.g., the $142^{+28}_{-16}\ \msun$ remnant observed from the gravitational wave event GW190521~\cite{Abbott:2020:PhRvL}). As a binary's total mass $\text{M}_{T}$ is associated with its gravitational-wave merger frequency, $f\sim \text{M}_{T}^{-1}$,  ground-based gravitational-wave detectors ($f\sim 10^1 - 10^3\ \text{Hz}$) are sensitive to the last milliseconds of merging systems with $100 \msun < \text{M}_{T} < 400 \msun$~\cite{LIGOScientificCollaboration:2015:CQGra, Martynov:2016:PhRvD, Moore_2014}, while space-based detectors ($f \sim 10^{-2} - 10^1\ \text{Hz}$) can study the full signals of merging systems with $10^4 \msun < \text{M}_{T} < 10^7 \msun$~\cite{ Moore_2014, Lu:2019:PhRvD}. Because of the short-duration of intermediate-mass gravitational-wave signals in ground-based detectors, data quality is critical for their detection. Gravitational-wave data is characterized by numerous non-stationary terrestrial artifacts called \textit{glitches}~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. Like signals from intermediate mass mergers, most glitches last for a fraction of a second, making them difficult to distinguish from astrophysical signals. These glitches can decrease the sensitivity of searches for binary black hole mergers with total masses $>55$~\msun ~\cite{pycbc_short_duration_transients}.

Although a significant fraction of the glitches can be identified by testing them for coherence amongst two or more detectors and performing matched-filtering, these methods are insufficient to identify all glitches~\cite{ pycbc_short_duration_transients, pe_with_glitch, blip_glitches}. One method to discriminate more glitches while searching for CBC is the Bayesian odds~\cite{bci, kanner2016leveraging, BCR1, BCR2, bcr_gw151216, bayesian_odds}. \et{Need sentence here saying we are inspired by the Bayesian odds to develop a ranking statistic that incorporates some of the ideas used in the odds.} In this paper, we rank O2's candidate gravitational-wave signals from $55-500\ \msun$ (\et{lab frame}) total mass systems using a ranking statistic called the Bayesian Coherence Ratio \bcr~\cite{BCR1}. \et{I deleted some text here going into detail about the BCR; save for method section.} For each candidate, we calculate the probability of astrophysical origin, \pastrobcr and compare to candidate events reported by other authors including the LIGO-Virgo-KAGRA (LVK) collaboration in \GWTC~\cite{GWTC1}, the \pycbc-team~\cite{pycbc_code, pycbc_og0, pycbc_og1, pycbc_og2, pycbc_og3, pycbc_og4, pycbc_og5, pycbc_og6, pycbc_single_det, pycbc_ogc_2}, by the Institute of Advanced study's team (\IAS)~\cite{IAS0, IAS1, IAS2}, or by \citet{bayesian_odds}. \et{Deleted several sentences here about innovations that I think are in the weeds.}

We find that (a) high-mass events reported in the \GWTC, including GW170729 (the heaviest event in \GWTC) are very statistically significant $\pastrobcr>0.9$; (b) three out of the eight \IAS events and candidates have $\pastrobcr>0.5$, corroborating \IAS's detection claims for GW170304, GW170727, and GW170817A; and that (c) our ranking statistic does not identify any new intermediate-mass black holes, but does identify an unreported marginal stellar-mass binary black hole candidate, 170222 with $\pastrobcr\sim0.5$. %\avi{Need to doublecheck that this wasnt in the PyCBC sub-threshold paper}.

@@

The standard framework to identify CBC gravitational-wave signals in data is to quantify the significance of candidates with null-hypothesis significance testing~\cite{GWTC1, GWTC2}. In this framework, the candidates' ranking statistic is compared against a background distribution. The independent matched-filter searches, e.g., \pycbc~\cite{pycbc_og4}, \spiir~\cite{spiir} and \gstlal~\cite{sachdev2019gstlal}, and Coherent WaveBurst~\cite{cwb} used by LVK to search for signals in gravitational-wave data all use ranking statistics in such a manner~\cite{GWTC1}. Both \pycbc and \gstlal's ranking statistic incorporate information about the relative likelihood that the data contains a coherent signal versus noise. In contrast, \cwb's ranking statistic uses the information of coherent energy present in the network of detectors~\cite{GWTC1}.

Bayesian inference offers an alternative means to rank the significance of candidate events by computing the odds that the data contain a transient gravitational-wave signal versus instrumental glitches~\cite{BCR1}. This method relies on accurate models for the signal and glitch morphologies~\cite{BCR1}. In principle, Bayesian odds is the optimal method for hypothesis testing~\cite{BCR2}. Much of its power comes from the Bayesian evidence, the likelihood of the data given a hypothesis. However, the evidence is not used in current matched filter searches. Here, we explore a hybrid frequentist/Bayesian ranking statistic that makes use of the Bayesian evidence. We compute the Bayesian evidence under the assumption that they either contain a coherent gravitational-wave signal, noise, or a glitch. However, instead of computing true Bayesian odds, we use the evidences as a ranking statistic.
We form a bootstrapped distribution of the evidence for simulated foreground and background events in order to form a frequentist ranking statistic.

\subsection{Formalism}

@@

When $\mathcal{H}_S$ and $\mathcal{H}_I$ are precisely described and the prior-odds represent our true beliefs, the \bcrodds is a Bayesian odds-ratio. As an odds-ratio, the \bcrodds is the optimal discriminator between coherent signals and incoherent instrumental features. However, as the priors-odds are unknown, we \textit{tune} values for prior-odds, $\hat{\pi}^S$ and $\hat{\pi}^G$, to estimate \bcrodds with a ranking statistic, \bcr, given by
\et{I would remove everything in this subsection before this point. Save it for your dissertation if you want, but I don't think we need to review Bayesian basics here. Instead, just begin by saying that we implement the BCR, which is given by Eq.~\ref{eq:rho_bcr} with suitable references to the previous BCR papers. The motivation for the BCR can be summarised in a single sentence.}
\begin{equation}
\label{eq:rho_bcr}
\bcr = \frac{\hat{\pi}^S Z^S}{\prod\limits^D_{i=1} \ [\hat{\pi}^G Z^G_i + (1-\hat{\pi}^G)Z^N_i]}\ .
\end{equation}
\et{Rewrite the remaining paragraphs to explain that how the BCR relates to the odds with just a few sentences, avoiding equations.}
In the limit where the estimated prior-odds equal the true prior-odds, $\bcr\to\bcrodds$. However, as we are uncertain what the true prior-odds are, it is invalid to use the \bcr as an odds-ratio to make an informed decision about whether a candidate is from an astrophysical or terrestrial source. Instead of interpreting the \bcr as a Bayesian odds-ratio, it can be used as a ranking statistic. Using the \bcr as a ranking statistic we can obtain a frequentist significance of a candidate \bcr-value measured against a background \bcr distribution.

@@

\subsection{Estimation of astrophysical signal probability}
\et{Above, I propose to cut the previous subsection way back. I think this subsection can be cut way back too and combined with the previous subsection. Once we have finished talking about $\rho_\text{BCR}$, we can just state that we use time-shifted data to generate distributions of $\rho_\text{BCR}$ for noise, which we use to calculate a single-event false alarm probability = the fraction of triggers with a $\rho_\text{BCR}$ at least that big. (The thing we are currently calling $f$ is a single-event false alarm probability, so we should call it that and give it a better variable name like $p_1$.) Then comment that we consider $N$ triggers so we also calculate a false-alarm probability with trial factors, which is given by the current Eq.~6. (I suggest you call the false alarm probability with trial factors $p_N$ to avoid confusion.) The text should say how $N$ is determined. I agree that $p_\text{astro} = 1-p_N$ is too sloppy to do in a paper. (The reason we are using FAP is because there was not appetite to calculate the distribution of $\rho_\text{BCR}$ for a plausible signal distribution, which is necessary to get an actual $p_\text{astro}$.) At this late stage, I think the best thing to do is just say that we are going to report $1-p_N$ alongside $p_\text{astro}$ values even though they are apples and oranges. While it's sloppy to compare them directly, they \textit{are} both indicators of significance. Moreover, different groups make different assumptions about the signal distribution, and so the different $p_\text{astro}$ values are sort of already different fruit; see Shanika's paper. We can explain all this and say we're going to put everything into one table anyway. However, the least bad thing to do is to report $p_N$ without calling it $p_\text{astro}$. I think most referees can live with that.}
Candidate \bcr-values are either statistically insignificant compared to the background \bcr distribution, implying the candidate is more probable to be an incoherent instrumental feature (the $\mathcal{H}_I$ null-hypothesis), or statistically significant relative to the background distribution, indicating the possible presence of an astrophysical signal (the $\mathcal{H}_S$ hypothesis). A false alarm probability with trial factors, \fap, for the candidate \bcr-value can quantify the significance. The \fap for a \bcr-value is the probability that a candidate originating from a non-astrophysical source can be falsely identified as a signal once in $N$ trials, and is given by 
\begin{equation}
    \fap = 1 - (1-f)^N \ ,
\end{equation}
where $f$ is the probability of observing a background $\bcr'$ greater than or equal to the candidate \bcr,
\begin{equation}
    f = \frac{\text{Count of } \bcr' \leq \bcr}{\text{Count of } \bcr'} \ .
\end{equation}

@@

\subsection{Background estimation}

To rank candidate event \bcr values we compute a background \bcr distribution. We use time-shifted background data from \pycbc~\cite{pycbc_og4}. \et{Add sentence here explaining what time-shifting is and why it is used.} The output of \pycbc's search is a list of times and their corresponding \pycbc ranking statistic \pycbcstat values. Whenever a local maximum of $\pycbcstat > \snr_\text{T}$, where $\snr_\text{T}$ is some predetermined threshold value, the \pycbc search pipeline produces a single-detector \textit{trigger} associated with the detector and time $t_c$ where the apparent signal in the data has its merger~\cite{pycbc_og6}.
For each \pycbc background trigger, we calculate \bcr.
% \avi{Should I mention that we need \pycbc  as a pre-selection step to computing? }
\et{I think this section can be cut way back too. (I've deleted a bunch of stuff w track changes.) I don't think we need to get into the three types of triggers; just focus on background triggers. I would move this discussion into a single subsection along with IIA and IIB. I would cover this stuff \textit{before}getting into FAP calculation. That way, we already have introduced the distribution of $\rho_\text{BCR}$; it will be more clear what we are talking about.}

@@

The most computationally intensive step during Bayesian inference is evaluating the likelihood $\mathcal{L}(d_i|\template)$. To accelerate our analysis, we use a likelihood that explicitly marginalizes over coalescence time, phase at coalescence, and luminosity distance (Eq.~80 from~\citet{intro_to_gw_bayes}).
\et{I think that IIIa and IIb should probably be shrunken and merged so that Section III does not have any subsections (IIC can be deleted; see below). The way to cut these sections down is to stick to just \textit{describing what was done} rather than trying to add a lot of explanation. (Explaining what PyCBC does and why is particularly excessive, IMO.) Just cut out anything that is not part of a recipe to reproduce the analysis.}

@@

\subsection{Assigning \pastro to candidate events}
After calculating the \bcr for the entire set of background and simulated triggers, we calculate the background and simulated \bcr probability distributions for each 2-week time-frame of O2 data. These distributions are used to `tune' prior-odd $\hat{\pi}^S$ and $\hat{\pi}^G$ values as described in Appendix~\ref{apdx:tuning-prior-odds}.
\et{This subsection discusses formalism (how the analysis works) and so it feels out of place in Section III, which is about the details of the implementation. I would delete this subsection entirely and just add a sentence or two to the formalism section explaining that the $\pi$'s are tuned emprically; see Appendix~\ref{apdx:tuning-prior-odds}.}

@@

\input{Data/results_table}
We analyze the $60,996$ background, $5,146$ simulated, and $25$ candidate triggers reported by \pycbc's search on the data from LIGO's second observing run, restricting our analysis to the triggers that fall within our mass-space as described in Section~\ref{sec:method}. We also analyze events and candidate events reported by \GWTC and the \IAS group (note: some of these were identified as candidates by the \pycbc search). Table~\ref{tab:results} summaries the \pastrobcr, along with the \pastro of other pipelines for comparison. 
\et{The previous text about what we analysed should go in Section III. This section is just for reporting the results.}

\et{Rewrite the following text w better explanation of FAP vs p-value.}
Although the various pipeline \pastro are not mathematically equivalent, by comparing pipeline \pastro values for a given candidate, we can compare how significant each pipeline deems various candidates. The $\hat{\pi}^S$ and $\hat{\pi}^G$ values utilized for each time-frame are reported in Appendix~\ref{apdx:alphabeta}.

@@

@@

%%%%%%---ACKNOWLEDGMENTS---%%%%%%%%%%%%
\begin{acknowledgments}
This work is supported by the Australian Research Council (ARC) Centre of Excellence CE170100004.
The author gratefully thank the \pycbc team for providing the gravitational-wave foreground, background and simulated triggers from \pycbc's search of O2's data. We also warmly thank Ian Harry and Thomas Dent for answering questions about the \pycbc search's data products.
